---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# gg1d

<!-- badges: start -->
[![R-CMD-check](https://github.com/selkamand/gg1d/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/selkamand/gg1d/actions/workflows/R-CMD-check.yaml)
[![CRAN status](https://www.r-pkg.org/badges/version/gg1d)](https://CRAN.R-project.org/package=gg1d)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![Codecov test coverage](https://codecov.io/gh/selkamand/gg1d/branch/main/graph/badge.svg)](https://app.codecov.io/gh/selkamand/gg1d?branch=main)
![GitHub Issues or Pull Requests](https://img.shields.io/github/issues-closed/selkamand/gg1d)
[![code size](https://img.shields.io/github/languages/code-size/selkamand/gg1d.svg)](https://github.com/selkamand/gg1d)
![GitHub last commit](https://img.shields.io/github/last-commit/selkamand/gg1d)
<!-- badges: end -->

Effortlessly visualize all columns in a data frame with vertically aligned plots and automatic plot selection based on variable type. Plots are fully interactive, and custom tooltips can be added.

**Why 1 dimensional plots?**

To understand trends in your data, especially correlative relationships between 2 or more features, it can be useful to densely stack visual representations of each feature vertically, regardless of data type. By unifying the $x$-axis across each plot, **gg1d** turns a series of 1D plots into an $n\text{-dimensional}$ visualization where $n = \text{number of columns in dataset}$. 
Note the key idea of gg1d is to 'preserve the individual.' **gg1d** does **NOT** plot distributions of properties, but rather each value of a feature for each subject/observation in the dataset.

gg1d can be used for exploratory data analysis (EDA) or to produce publication quality graphics summarizing a dataset.


## Installation

You can install the development version of gg1d from [GitHub](https://github.com/) with:


``` r
if (!require("remotes"))
    install.packages("remotes")

remotes::install_github("selkamand/gg1d")
```


## Quick Start

For examples of interactive gg1d plots see the [gg1d gallery](https://selkamand.github.io/gg1d/articles/gallery.html)

```{r, example, fig.width = 7, fig.height = 5}
# Load library
library(gg1d)

# Read data
path_gg1d <- system.file("example.csv", package = "gg1d")
df <- read.csv(path_gg1d, header = TRUE, na.strings = "")

# Plot data, sort by Glasses
gg1d(
  df,
  col_id = "ID",
  col_sort = "Glasses",
  interactive = FALSE,
  verbose = FALSE,
  options = gg1d_options(legend_nrow = 2)
)
```

## Customise Colours

Customise colours by supplying a named list to the `palettes` argument

```{r customise_colours, fig.width = 7, fig.height = 5}
gg1d(
  df,
  col_id = "ID",
  col_sort = "Glasses",
  palettes = list("EyeColour" = c(
    Brown = "rosybrown4",
    Blue = "steelblue",
    Green = "seagreen"
  )),
  interactive = FALSE,
  verbose = FALSE,
  options = gg1d_options(legend_nrow = 2)
)
```

## Community Contributions

All types of contributions are encouraged and valued. See our [guide to
community contributions](CONTRIBUTING.md) for different ways to help.
