---
title: "advanced_interactivity"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{tooltips}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(gg1d)
```

## Adding Custom Tooltips

Custom tooltips make your visualizations more interactive and insightful. You can easily add tooltips to any column in your dataset by creating a new column with the suffix `_tooltip`

### Example: Adding Tooltips to the Magpies Barplot

Let’s use the lazy_birdwatcher dataset to add custom tooltips for the Magpies column. By adding a Magpies_tooltip column, we can provide context for the values, such as highlighting when there are more than three magpies.

```{r fig.height=6, message=FALSE}
# No Custom Tooltip
gg1d(lazy_birdwatcher)
```

```{r}
# Add a custom tooltip column
lazy_birdwatcher["Magpies_tooltip"] <- ifelse(
  lazy_birdwatcher[["Magpies"]] > 3, 
  yes = "More than 3 magpies!", 
  no = paste0("How boring. Too few magpies (", lazy_birdwatcher[["Magpies"]], ")")
)

# Generate a gg1d plot with the custom tooltip
gg1d(lazy_birdwatcher)
```


## Cross-linking gg1d plots with other visualisatios

gg1d plots can be cross-linked with visualisations produced by custom scripts or other packages.  

In this example we will produce both gg1d visualisation and a umap summarising the `palmer_penguins` and render them together so they can be coexplored.


### Workflow

1. Create your gg1d plot. Ensure you set `col_id` to embed a data-identifier required for cross-linking plots (in our case, we will use the unique name given to the penguins). Also set `interactive = FALSE` to force gg1 to returns a ggplot object instead of a htmlwidget (we'll add the interactivity back in after we compose all our plots). For example, lets 

2. Create the ggplot you want to co-explore with, adding latent interactivity using the ggiraph package (e.g. use `geom_point_interactive` instead of `geom_point` and add a `data_id` aesthetic set to the same column as col_id)

3. Compose the two ggplots together using patchwork

4. Make the plots interactive using the ggiraph package


### Example Code 

```{r fig.height = 6}
library(uwot)
library(ggplot2)
library(patchwork)
library(ggiraph)


# Prepare Data & Colour Schemes
penguins <- na.omit(read.csv(system.file("penguins.csv", package = "gg1d")))
palette_species = c(Chinstrap = "#C55BCC", Adelie = "#FF7F02", Gentoo = "#047476")

# Create gg1d plot
gg1d_penguins <- gg1d(
  penguins, 
  col_id = "name", 
  interactive = FALSE, 
  col_sort = "species", 
  palettes = list(species = palette_species),
  options = gg1d_options(relative_height_numeric = 1.2, show_legend = FALSE)
  )

# Perform the umap
umap_mx <- umap(penguins, scale = "scale", n_neighbors = 70)
df_umap <- as.data.frame(umap_mx)
colnames(df_umap) <- c("UMAP_1", "UMAP_2")
df_umap[["name"]] <- penguins[["name"]]
df_umap[["species"]] <- penguins[["species"]]

# Create the umap plot 
gg_umap <- ggplot(df_umap, aes(x = UMAP_1, y = UMAP_2, colour = species, data_id = name)) +
  geom_point_interactive() + 
  scale_color_manual(values = palette_species) +
  theme_bw()

# Combine the gg1d plot and the umap using patchwork
combined_plots <- free(gg_umap, type = "label") / free(gg1d_penguins, side = "tb")

# Make interactive
interactive_multiplot <- girafe(ggobj = combined_plots, height_svg = 6, width_svg = 10, options )

# Add some settings to choose how to make combined plots interactive
interactive_multiplot <- girafe_options(x = interactive_multiplot,
    opts_selection(type = "multiple", only_shiny = FALSE, css = "opacity: 1"),
    opts_selection_inv(css = "opacity: 0.12")
    )

interactive_multiplot
```
Create

