---
title: "Gallery"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Gallery}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width =  13,
  fig.height = 9 
)
```

```{r setup}
library(gg1d)
library(datarium)
library(palmerpenguins)
```

## Examples with real datasets

### Chicken Weights by Feed

Horsebean feed consistently produces the lowest chicken weights.

```{r message=FALSE, fig.height=3}
gg1d(
  chickwts, 
  show_legend = TRUE, 
  legend_position = "bottom", 
  legend_nrow = 1, 
  legend_text_size = 18, 
  col_sort = "weight")
```

### Iris

The setosa species have drastically smaller petals then other iris species.

```{r message=FALSE, fig.height=7}
gg1d(iris, show_legend = TRUE)
```

### Titanic Survival by Class

Most deaths in the titanic were male adults.

```{r,  message=FALSE, fig.height=4}
gg1d(
  titanic.raw,
  palettes = list(
    Sex = c(Male = "#0072B2", Female = "#CC79A7"),
    Survived = c(Yes = "#32A02D", No ="#E31A1C"),
    Class = c("Crew" =  "#F9F7EAFF", "1st" ="#006E37FF", "2nd" = "#8DB580FF", "3rd" =  "#D9E5BDFF"),
    Age = c(Child = "#35B779FF",  Adult = "#440154FF")
    ), show_legend = TRUE
  )
```

### Palmer Penguins

Gentoo penguins from Biscoe Island have the shallowest bills depths.

```{r, fig.height=7}

library(palmerpenguins)

# Drop Year Column
penguins$year <- NULL

gg1d(
  penguins, 
  relative_height_numeric = 1.2, 
  show_legend = TRUE, 
  palettes = list(
    species = c(Chinstrap = "#C55BCC", Adelie = "#FF7F02", Gentoo = "#047476"),
    sex = c(male = "#0072B2", female = "#CC79A7"),
    island = c(Biscoe = "#E69F00", Dream = "#56B4E9", Torgersen = "#009E73")
    )
)
```

## Artificial Data

### Inbuilt gg1d example

```{r message=FALSE, fig.height = 6}
# Read data
path_gg1d <- system.file("example.csv", package = "gg1d")
df <- read.csv(path_gg1d, header = TRUE, na.strings = "")

# Plot data, sort by Glasses
gg1d(
  df, 
  col_id = "ID", 
  col_sort = "Glasses", 
  legend_nrow = 2
)
```

### Lazy Birdwatcher

The simulated "Lazy Birdwatcher" dataset records magpie observations by Robert and Catherine. Robert doesn't birdwatch on weekends, resulting in missing data for him on those days. This pattern of missingness, dependent on both person and day, is hard to detect using one-dimensional EDA tools like skimr or two-dimensional tools like ggpairs from the GGally package, as they don't capture multi-variable interactions contributing to the missingness.

```{r message=FALSE, fig.height = 4}
# Read Data
path_lazy_birdwatcher <- system.file("birdwatching.csv", package = "gg1d")
df_lazy_birdatcher <- read.csv(path_lazy_birdwatcher)

# Plot Dataset by Number of Magpies Observed
gg1d(
  df_lazy_birdatcher,
  col_sort = "Magpies", 
  show_legend = TRUE,
  palettes = list(
    Birdwatcher = c(Robert = "#E69F00", Catherine = "#999999"),
    Day = c(Weekday = "#999999" ,Weekend = "#009E73")
  )
)
```
